---
title: "Project 1"
author: "Sanskriti Purohit & Caleb Woo"
date: "2022-10-13"
output: pdf_document
---

```{r chunk-options}
knitr::opts_chunk$set(message = F, warning = F)
```

```{r libraries-data}
library(tidyverse)
library(dplyr)
library(chron)
library(lubridate)
library(ggcorrplot)
library(RColorBrewer)
library(scales)
library(kableExtra)
library(gridExtra)
library(factoextra)
data_net <- read.csv("data/sonoma-data-net.csv")
data_log <- read.csv("data/sonoma-data-log.csv")
dates <- read_csv("data/dates.csv")
dates <- dates %>%
  select(-c("...1"))
```
## 1\newline
### 1.1 Summary of the Paper \newline
The paper, "A Macroscope in the Redwoods" highlights the potential of the wireless sensor network macroscope in the field of science, particularly its use in observing and monitoring instances that vary over time and space. It also provides a way to analyse the complex data obtained through these sensors and draws a few conclusions about the dynamics of the microclimate around a coastal redwood tree. Therefore, establishing and achieving its motivation of devising a base model for future researches to observe what they weren't able to measure at the time. The redwood tree, proven to have a substantial variation in its microclimate over both space and time, was thus chosen to be the center of this research.  \newline
During the time frame of a month, the collected data, measurements of temperature, humidity, incident PAR and reflected PAR, was transferred over a network and also logged through a distributed data logger  contained in the suite of sensors placed on the redwood tree, such that each sensor maintained a certain minimum and maximum distance from the ground level, the trunk and other sensors. \newline
By projecting the sensor values onto time and then onto space, the researchers discovered a much higher variation over time than in space which informed them to remove temporal variation and focus on spatial variation.The researchers determined that PAR and temperature correlate with solar movement, although reflected PAR is more noisy because the absolute readings are lower.Humidity, on the other hand, was shown to change considerably over a wide range prior to sunrise and after sunrise spatial variability remains high suggesting that the foliage of the tree has a significant effect on humidity across different heights of the tree. \newline
Thus, multi-dimensional analysis of the data paints a detailed picture of the spatiotemporal variance, previously validated by biologists. Further, it shines light on the shortcomings of the existing wireless sensor networks. \newline
### 1.2 Summary of Data Collection \newline
The data for the study was collected over a duration of approximately 44 days (Tuesday, April 27, 2004 at 5:10pm to Thursday, June 10, 2004 at 2:00pm), where each sensor was sampled every 5 minutes. The sensors were deployed on the west side of the tree, at least 15m above the ground level, reaching a maximum distance of 70m while maintaining a 0.1-1 meter distance from the tree trunk. This was to ensure the capture of the variation caused by the foliage. There was also a 2 meter distance between each node, which proved to have a significant impact on the results obtained. \newline
The package was designed to ensure protection of the components of the node from various natural phenomenon such as rain, dirt, wind e.t.c. while enabling them to have the required sensitivity to measure the variables of interest.
The researchers utilized the TinyDB data collection framework to collect the data into table format and represent each sensor as a column and each row as a reading taken at a particular time with additional columns for node ID, sample number, and sample reception time. Further the hardware and the software was calibrated twice to ensure optimum collection of the measurements of temperature, humidity, incident PAR and reflected PAR. Roof calibration was used on the PAR sensors to ensure that they were producing acceptable readings relative to a high-quality-well-established PAR sensor.
Two-point calibration for each humidity and temperature sensor was performed by placing sensors in a controllable weather chamber and cycling through various temperatures and humidity levels.
While calibration improved the mean temperature deviation from 0.35 degrees Celsius to 0.18 degrees Celsius and the mean humidity deviation from $1.24\%$ relative humidity to $0.60\%$ relative humidity, the absolute increase in accuracy was small enough to be of minimal benefit.\newline
The data was collected in two ways, through the transfer of the data to the gateway over a network and logging of the data at the site by the data logger installed in each node. \textbf{sonoma-data-log.csv} contains the data obtained from the logger while \textbf{sonoma-data-net.csv} contains that transferred over the network.

## 2.

### a)

Voltage

```{r initial-voltage}
data_net %>%
  ggplot(data=., aes(x=voltage)) +
  geom_histogram(binwidth=25) +
  labs(title="Network ADC Histogram", x="ADC", y="Count")

data_log %>%
  ggplot(data=., aes(x=voltage)) +
  geom_histogram(binwidth=0.25) +
  labs(title="Logs Voltage Histogram", x="Voltage", y="Count")
```

Voltage is inconsistent, link below for ADC to voltage conversion

  - http://www-db.ics.uci.edu/pages/research/quasar/MPR-MIB%20Series%20User%20Manual%207430-0021-06_A.pdf
  
  - page 23 (25 of pdf) has MICA2DOT conversion

```{r fixed-voltage}
data_net <- data_net %>%
  mutate(voltage = 0.6*(1024/voltage))

data_net %>%
  ggplot(data=., aes(x=voltage)) +
  geom_histogram(binwidth=0.25) +
  labs(title="Network Voltage Histogram", x="Voltage", y="Count")

data_log %>%
  ggplot(data=., aes(x=voltage)) +
  geom_histogram(binwidth=0.25) +
  labs(title="Logs Voltage Histogram", x="Voltage", y="Count")
```

```{r humidity}
data_net %>%
  filter(!is.na(humidity) & voltage <= 3 & voltage >= 2.4 & humidity < 105) %>%
  ggplot(data=., aes(x=humidity)) +
  geom_histogram(bins=20) +
  labs(title="Network Humidity Histogram", x="Humidity", y="Count")

data_log %>%
  filter(!is.na(humidity) & voltage <= 3 & voltage >= 2.4) %>%
  ggplot(data=., aes(x=humidity)) +
  geom_histogram(bins=20) +
  labs(title="Logs Humidity Histogram", x="Humidity", y="Count")
```
```{r adj-humidity}
data_net %>%
  filter(!is.na(humid_adj) & voltage <= 3 & voltage >= 2.4 & humid_adj < 105) %>%
  ggplot(data=., aes(x=humid_adj)) +
  geom_histogram(bins=20) +
  labs(title="Network Adjusted Humidity Histogram", x="Humidity", y="Count")

data_log %>%
  filter(!is.na(humid_adj) & voltage <= 3 & voltage >= 2.4) %>%
  ggplot(data=., aes(x=humid_adj)) +
  geom_histogram(bins=20) +
  labs(title="Logs Adjusted Humidity Histogram", x="Humidity", y="Count")
```
```{r temperature}
data_net %>%
  filter(!is.na(humid_temp) & voltage <= 3 & voltage >= 2.4 & humid_temp < 35) %>%
  ggplot(data=., aes(x=humid_temp)) +
  geom_histogram(bins=15) +
  labs(title="Network Temperature Histogram", x="Humidity", y="Count")

data_log %>%
  filter(!is.na(humid_temp) & voltage <= 3 & voltage >= 2.4) %>%
  ggplot(data=., aes(x=humid_temp)) +
  geom_histogram(bins=15) +
  labs(title="Logs Temperature Histogram", x="Humidity", y="Count")
```

Incident PAR
show histograms before conversion
don't filter yet

```{r initial-hamatop}
data_net %>%
  filter(!is.na(hamatop) & voltage <= 3 & voltage >= 2.4) %>%
  select(hamatop) %>%
  pull() %>%
  min()

data_log %>%
  filter(!is.na(hamatop) & voltage <= 3 & voltage >= 2.4) %>%
  select(hamatop) %>%
  pull() %>%
  min()

data_net %>%
  filter(!is.na(hamatop) & voltage <= 3 & voltage >= 2.4) %>%
  select(hamatop) %>%
  pull() %>%
  max()

data_log %>%
  filter(!is.na(hamatop) & voltage <= 3 & voltage >= 2.4) %>%
  select(hamatop) %>%
  pull() %>%
  max()
```

Hamatop is inconsistent, link below for Lux to PPFD conversion

  - https://www.apogeeinstruments.com/conversion-ppfd-to-lux/

```{r fixed-hamatop}
data_net <- data_net %>%
  mutate(hamatop=0.0185*hamatop)

data_log <- data_log %>%
  mutate(hamatop=0.0185*hamatop)

data_net %>%
  filter(!is.na(hamatop) & voltage <= 3 & voltage >= 2.4) %>%
  ggplot(data=., aes(x=hamatop)) +
  geom_histogram(bins=20) +
  labs(title="Network Incident PAR Histogram", x="Incident PAR", y="Count")

data_log %>%
  filter(!is.na(hamatop) & voltage <= 3 & voltage >= 2.4  & hamatop < 2200) %>%
  ggplot(data=., aes(x=hamatop)) +
  geom_histogram(bins=20) +
  labs(title="Logs Incident PAR Histogram", x="Incident PAR", y="Count")
```

Reflected PAR
show histograms before conversion

```{r initial-hamabot}
data_net %>%
  filter(!is.na(hamabot) & voltage <= 3 & voltage >= 2.4) %>%
  select(hamabot) %>%
  pull() %>%
  min()

data_log %>%
  filter(!is.na(hamabot) & voltage <= 3 & voltage >= 2.4) %>%
  select(hamabot) %>%
  pull() %>%
  min()

data_net %>%
  filter(!is.na(hamabot) & voltage <= 3 & voltage >= 2.4) %>%
  select(hamabot) %>%
  pull() %>%
  max()

data_log %>%
  filter(!is.na(hamabot) & voltage <= 3 & voltage >= 2.4) %>%
  select(hamabot) %>%
  pull() %>%
  max()
```

Fix hamabot like hamatop

```{r fixed-hamabot}
data_net <- data_net %>%
  mutate(hamabot=0.0185*hamabot)

data_log <- data_log %>%
  mutate(hamabot=0.0185*hamabot)

data_net %>%
  filter(!is.na(hamabot) & voltage <= 3 & voltage >= 2.4) %>%
  ggplot(data=., aes(x=hamabot)) +
  geom_histogram(bins=15) +
  labs(title="Network Reflected PAR Histogram", x="Reflected PAR", y="Count")

data_log %>%
  filter(!is.na(hamabot) & voltage <= 3 & voltage >= 2.4) %>%
  ggplot(data=., aes(x=hamabot)) +
  geom_histogram(bins=15) +
  labs(title="Logs Reflected PAR Histogram", x="Reflected PAR", y="Count")
```

### b)

```{r dates-merge}
data_both <- merge(data_net, data_log, by = c("epoch", "nodeid"), all.y = T, suffixes = c(".net", ".log"))
data_all <- merge(data_both, dates, by = "epoch", all.x = T)
```

```{r missing}
missing_data <-  data_all %>%
  select(ends_with(".log"), epochDates, epochDays) %>%
  filter_all(any_vars(is.na(.)))
  #filter(is.na(parent.log) | is.na(voltage.log) | is.na(depth.log) | is.na(humidity.log) | is.na(humid_temp.log) |
  #       is.na(humid_adj.log) | is.na(hamatop.log) | is.na(hamabot.log))

paste0("# of missing measurements: ", nrow(missing_data))

missing_data %>%
  select(epochDates) %>%
  summarise(min = min(epochDates),
            max = max(epochDates))
```

```{r remove-missing}
data_all <- data_all %>%
  filter(!is.na(parent.log) & !is.na(voltage.log) & !is.na(depth.log) & !is.na(humidity.log) & !is.na(humid_temp.log) &
         !is.na(humid_adj.log) & !is.na(hamatop.log) & !is.na(hamabot.log))
row.names(data_all) <- NULL
```

### c)

```{r locations}
locations <- read_csv("data/locations.csv") %>%
  select(-c("...1"))
```

```{r locations-merge}
data_all <- merge(data_all, locations, by = "nodeid", all.x = T)
```

The network data and logs data have identical columns in this merged data frame except for voltage. This must be because network voltage data had to be converted from ADC so the conversion may not be exact. Since network data is missing for some rows, logs data is not missing for any rows, and logs data has identical values as network data (other than voltage but logs voltage is more accurate anyways), we can remove network data columns from the merged data frame. In addition, result_time and epochDates represent the same information with epochDates from sonoma-dates being more accurate, so I will also remove the result_time column from the logs data.

```{r remove-network-data}
data_all <- data_all %>%
  select(-ends_with(".net"), -result_time.log)
names(data_all)[3:11] <- str_replace(names(data_all)[3:11], ".log", "")
```

There are 16 total variables in my complete merged data frame that includes network and logs data, dates data, and locations data.

```{r remove-missing-locations}
data_all %>%
  filter(is.na(Height) & is.na(Direc) & is.na(Dist) & is.na(Tree)) %>%
  nrow()
data_all <- data_all %>%
  filter(!is.na(Height) & !is.na(Direc) & !is.na(Dist) & !is.na(Tree))
```

6091 missing location observations.

### d)

Quantiles before removal, quantiles and histograms after removal

Since sensors with low batteries were not accurate, we removed all data with voltages greater than 3 and less than 2.4 like Tolle et al. This removed 26011 observations. Below are the quantiles and histograms of each variable after removing the voltages and before removing any other outliers.

```{r voltage-outliers}
data_all %>%
  filter(voltage > 3 | voltage < 2.4) %>%
  nrow()
data_all <- data_all %>%
  filter(voltage <= 3 & voltage >= 2.4)
```

```{r humid_adj-removal}
quantile(data_all$humid_adj)

data_all %>%
  ggplot(., aes(x=humid_adj)) +
  geom_histogram()
```

The quantiles in humid_adj more closely matches that of Tolle et al. compared to the quantiles of humidity so we should use humid_adj in place of humidity. This allows us to remove 1 more variable so we have a total of 15 variables. There do not seem to be any other outliers in humid_adj.

```{r hamatop-pre-removal}
quantile(data_all$hamatop)

data_all %>%
  ggplot(., aes(x=hamatop)) +
  geom_histogram()
```

```{r high-hamatop-nodes}
data_all %>%
  filter(hamatop > 2154) %>%
  select(nodeid) %>%
  unique()
```

There seem to be some abnormally high values of incident PAR and when we look for the nodes that gave a reading higher than the max incident PAR reading of 2154 given in the paper, we see that node 40 is solely responsible for all those readings. So we will remove all observations with node 40. This removes 98 observations. 271,776 observations left.

```{r hamabot-pre-removal}
quantile(data_all$hamabot)

data_all %>%
  ggplot(., aes(x=hamabot)) +
  geom_histogram()
```

The quantiles of reflected PAR match up well with that of Tolle et al. so there does not seem to be any other outliers in reflected PAR.

Below are the quantiles and histograms of each variable after removing both the voltages and observations with node 40. 

```{r remove-node-40}
data_all %>%
  filter(nodeid == 40) %>%
  nrow()
data_all <- data_all %>%
  filter(nodeid != 40)
```

```{r humid-adj-post-removal}
quantile(data_all$humid_adj)

data_all %>%
  ggplot(., aes(x=humid_adj)) +
  geom_histogram()
```

```{r temp-post-removal}
quantile(data_all$humid_temp)

data_all %>%
  ggplot(., aes(x=humid_temp)) +
  geom_histogram()
```

```{r hamatop-post-removal}
quantile(data_all$hamatop)

data_all %>%
  ggplot(., aes(x=hamatop)) +
  geom_histogram()
```

```{r hamabot-post-removal}
quantile(data_all$hamabot)

data_all %>%
  ggplot(., aes(x=hamabot)) +
  geom_histogram()
```

All quantiles and histograms of humidity, temperature, incident PAR, and reflected PAR match up well with the quantiles of Tolle et al. after removing observations with voltages greater than 3 and less than 2.4 and removing observations with node 40.

```{r write-csv-data-all}
data_all %>%
  filter_all(any_vars(is.na(.))) %>%
  nrow()

#write.csv(data_all, "data/data-all-dates-locations.csv")
```

```{r}
#Checking for further outliers
data_all %>%
  ggplot(., aes(x=voltage))+geom_boxplot()
data_all %>%
  ggplot(., aes(x=humid_temp))+geom_boxplot()
```

## 3.

### a)
```{r}
data_all = read.csv("data/data-all-dates-locations.csv")
data_all %>%
  select(epochDates) %>%
  summarise(
    min = min(epochDates),
    max = max(epochDates)
  )

data_all$Date = as.Date(data_all$epochDates)
data_all$time = format(as.POSIXct(data_all$epochDates), format = "%H:%M")
data_all$dayofweek = wday(data_all$epochDates)
data_all$hour = format(as.POSIXct(data_all$epochDates), format = "%H")
data_all$hour = as.numeric(data_all$hour)
# data_all$Height = as.numeric(data_all$Height) #Check this
#Reason for choosing the hours for Dark and Light
data_all %>% 
  ggplot(aes(x=hour,y=hamatop)) + geom_point()
data_all$Sunlight = case_when(
  as.numeric(data_all$hour) <= 5 | as.numeric(data_all$hour) >= 20  ~ "Dark",
  as.numeric(data_all$hour) < 20 ~ "Light"
)
count(data_all,Sunlight)

date_count = data_all %>% 
  count(Date)
date_count$n = date_count$n*100/sum(date_count$n)
p=0
for(i in 1:length(date_count$n)){
  p = p + date_count$n[i]
  date_count$n[i] = p
}
date_count$n
data_all = data_all %>%
  filter(Date > '2004-04-27' & Date < '2004-05-27')
data_all %>%
  select(epochDates) %>%
  summarise(
    min = min(epochDates),
    max = max(epochDates)
  )
count(data_all,Sunlight)

#Add more?
data_all %>% 
  ggplot(aes(x=humid_temp, y=humid_adj, col = Sunlight)) + geom_point(alpha=0.05)
data_all %>%
  ggplot(aes(x=hour,y=hamatop)) + geom_point()
data_all %>%
  ggplot(aes(x=humid_temp,y=hamatop)) + geom_point()
data_all %>%
  ggplot(aes(x=humid_adj,y=hamatop)) + geom_point()
data_all %>% 
  ggplot(aes(x=humid_adj,y=hamabot, col = Sunlight)) + geom_point(alpha=0.05)
data_all %>% 
  ggplot(aes(x=humid_temp,y=hamabot, col=Sunlight)) + geom_point(alpha=0.05)
#Abnormal values of reflected ray at odd hours at higher levels of heights.
data_all %>% 
  ggplot(aes(x=hour,y=hamabot)) + geom_point()
```
## 3
### b)
```{r}
corr=cor(data_all[,unlist(lapply(data_all,is.numeric))],y=data_all$hamatop, method=c("pearson","kendall","spearman"))
ggcorrplot(corr,method="square")
```

### c)
```{r}
data_all %>%
  group_by(Date,Height)%>%
  summarise(avg_hamatop = mean(hamatop),.groups="drop") %>%
  ggplot(aes(x=Date,y=avg_hamatop,col=Height,group=Height))+ geom_line()+scale_x_date(breaks="4 days",date_labels = "%B %d")
data_all %>%
  group_by(Date,Height)%>%
  summarise(avg_hamabot = mean(hamabot),.groups="drop") %>%
  ggplot(aes(x=Date,y=avg_hamabot,col=Height,group=Height)) + geom_line()+scale_x_date(breaks="4 days",date_labels = "%B %d")
data_all %>%
  group_by(Date,Height)%>%
  summarise(avg_humid = mean(humid_adj),.groups="drop") %>%
  ggplot(aes(x=Date,y=avg_humid,col=Height,group=Height)) + geom_line()+scale_x_date(breaks="4 days",date_labels = "%B %d")
data_all %>%
  group_by(Date,Height)%>%
  summarise(avg_temp = mean(humid_temp),.groups="drop") %>%
  ggplot(aes(x=Date,y=avg_temp,col=Height,group=Height)) + geom_line()+scale_x_date(breaks="4 days",date_labels = "%B %d")


data_all %>%
  group_by(dayofweek,Height)%>%
  summarise(avg_hamatop = mean(hamatop),.groups="drop") %>%
  ggplot(aes(x=dayofweek,y=avg_hamatop,col=Height,group=Height))+ geom_line()
data_all %>%
  group_by(dayofweek,Height)%>%
  summarise(avg_hamabot = mean(hamabot),.groups="drop") %>%
  ggplot(aes(x=dayofweek,y=avg_hamabot,col=Height,group=Height)) + geom_line()
data_all %>%
  group_by(dayofweek,Height)%>%
  summarise(avg_humid = mean(humid_adj),.groups="drop") %>%
  ggplot(aes(x=dayofweek,y=avg_humid,col=Height,group=Height)) + geom_line()
data_all %>%
  group_by(dayofweek,Height)%>%
  summarise(avg_temp = mean(humid_temp),.groups="drop") %>%
  ggplot(aes(x=dayofweek,y=avg_temp,col=Height,group=Height)) + geom_line()


data_all %>%
  group_by(hour,Height)%>%
  summarise(avg_hamatop = mean(hamatop),.groups="drop") %>%
  ggplot(aes(x=hour,y=avg_hamatop,col=Height,group=Height))+ geom_line()
data_all %>%
  group_by(hour,Height)%>%
  summarise(avg_hamabot = mean(hamabot),.groups="drop") %>%
  ggplot(aes(x=hour,y=avg_hamabot,col=Height,group=Height)) + geom_line()
data_all %>%
  group_by(hour,Height)%>%
  summarise(avg_humid = mean(humid_adj),.groups="drop") %>%
  ggplot(aes(x=hour,y=avg_humid,col=Height,group=Height)) + geom_line()
data_all %>%
  group_by(hour,Height)%>%
  summarise(avg_temp = mean(humid_temp),.groups="drop") %>%
  ggplot(aes(x=hour,y=avg_temp,col=Height,group=Height)) + geom_line()
```
### d)
```{r}
names(data_all)
cols_name = c("hamatop","hamabot","Height","humid_temp","humid_adj")
data_filtered = data_all |>
  select(all_of(cols_name)) |>
  data.frame()
pca_prcomp = prcomp(data_filtered, scale. = TRUE)
summary(pca_prcomp)
eigenvalues = pca_prcomp$sdev**2
plot(1:length(eigenvalues), eigenvalues/sum(eigenvalues), type = "l")
points(1:length(eigenvalues), eigenvalues/sum(eigenvalues))
```

## 4
Through the k-means clustering of the Principal Components, it can be observed that while plotting the components that explain a high variability of the data, humidity and temperature appear in opposite directions, hence confirming their negative correlation with each other. \newline
Also, when we consider the 3 clusters in the plot of PC1 against PC2, which cumulatively explain approx. $72\%$ of our data, we can observe that the darkest cluster (represented by 1.0) represents values that contain high values of height, incident PAR and reflected PAR and relatively high values of temperature, leading to lower values of relative humidity. The second cluster (represented by 2.0) -> Lower values of height with higher temperature, higher incident PAR, higher reflect PAR and lower humidity. Third cluster (represented by 3.0) -> lower values of height, higher humidity, lower temperature, lower incident and reflected PAR.
```{r k means PCA}
#Try grouping and checking over the week
comp = data.frame(pca_prcomp$x)
wss = function(k) {
  kmeans(comp,algorithm="Lloyd", k, nstart = 10, iter.max=200 )$tot.withinss
}

k.values <- 1:5
wss_values <- map_dbl(k.values, wss)
plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")
# 3?
kmeans_pca = kmeans(comp,3)
fviz_pca(pca_prcomp, axes = c(1, 2),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(1, 3),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(1, 4),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(1, 5),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(2, 3),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(2, 4),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(2, 5),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(3, 4),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(3, 5),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
fviz_pca(pca_prcomp, axes = c(4, 5),geom=c("point"),col.ind = kmeans_pca$clust,col.var="red",alpha.ind=0.25)
```

```{r may-count}
data_all <- read_csv("data/data-all-dates-locations.csv") %>%
  select(-c("...1"))
data_all %>%
  filter(as.Date(epochDates) >= as.Date("2004-05-01") & as.Date(epochDates) <= as.Date("2004-05-31")) %>%
  nrow()
```

78.2% of data is within May, 2004

```{r count-Times}
data_all_sun <- read_csv("data/may-data-sun.csv")
data_all_sun$Time <- factor(data_all_sun$Time, levels = c("Early Night", "Late Night", "Sunrise", "Early Morning", "Late Morning", "Early Noon", "Late Noon", "Sunset"))
data_all_sun %>%
  count(Time)
```

### Time vs Variable Boxplots

```{r boxplots}
ggplot(data_all_sun, aes(x=Time, y=hamatop, group=Time)) +
  geom_boxplot()

ggplot(data_all_sun, aes(x=Time, y=humid_temp, group=Time)) +
  geom_boxplot()

ggplot(data_all_sun, aes(x=Time, y=humid_adj, group=Time)) +
  geom_boxplot()
```

### Clustering

```{r clust-initial-plot}
data_all_sun %>%
  filter(Time == "Early Night" | Time == "Early Noon") %>%
  ggplot(., aes(x=humid_adj, y=humid_temp, color=Time)) +
  geom_point() +
  labs(title="Early Night vs Early Noon", x="Humidity", y="Temperature")
```

```{r clustering-1}
library(mclust)
set.seed(521)

gmm1 <- data_all_sun %>%
  filter(Time == "Early Night" | Time == "Early Noon") %>%
  select(humid_adj, humid_temp) %>%
  Mclust(., G=5, modelNames="VVV", prior=priorControl())

clusters_data_1 <- data_all_sun %>%
  filter(Time == "Early Night" | Time == "Early Noon") %>%
  select(humid_adj, humid_temp) %>%
  cbind(., gmm1$classification)

names(clusters_data_1)[3] <- "Cluster"
clusters_data_1$Cluster <- as.factor(clusters_data_1$Cluster)

ggplot(clusters_data_1, aes(x=humid_adj, y=humid_temp, color=Cluster)) +
  geom_point() +
  labs(title="GMM Clustering", x="Humidity", y="Temperature")
```

Cluster 1 = Early Noon, Cluster 2 = Early Night, Cluster 3 = Early Night, Cluster 4 = Early Noon, Cluster 5 = Early Night

```{r cluster-accuracy}
clusters_real_data <- data_all_sun %>%
  filter(Time == "Early Night" | Time == "Early Noon") %>%
  select(humid_adj, humid_temp, Time, Hour, Day, Height) %>%
  cbind(., gmm1$classification)
names(clusters_real_data)[7] <- "Cluster"
clusters_real_data$Cluster <- as.factor(clusters_real_data$Cluster)

acc1 <- clusters_real_data %>%
  filter(Cluster == 1) %>%
  count(Time) %>%
  summarise(X1 = 100*n[2]/sum(n))

acc2 <- clusters_real_data %>%
  filter(Cluster == 2) %>%
  count(Time) %>%
  summarise(X2 = 100*n[1]/sum(n))

acc3 <- clusters_real_data %>%
  filter(Cluster == 3) %>%
  count(Time) %>%
  summarise(X3 = 100*n[1]/sum(n))

acc4 <- clusters_real_data %>%
  filter(Cluster == 4) %>%
  count(Time) %>%
  summarise(X4 = 100*n[2]/sum(n))

acc5 <- clusters_real_data %>%
  filter(Cluster == 5) %>%
  count(Time) %>%
  summarise(X5 = 100*n[1]/sum(n))
```

Cluster 1 = Early Noon, Cluster 2 = Early Night, Cluster 3 = Early Night, Cluster 4 = Early Noon, Cluster 5 = Early Night

```{r gmm-means}
gmm_means <- rbind(cbind(acc1, acc2, acc3, acc4, acc5), data.frame(gmm1$parameters$mean))
names(gmm_means) <- c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5")
rownames(gmm_means) <- c("Accuracy %", "Humidity Mean", "Temperature Mean")
gmm_means %>%
  kbl(digits = 3, caption = "GMM Accuracy and Cluster Means") %>%
  kable_styling(latex_options = "HOLD_position", full_width = F)
```

Clusters 4 and 5 both have low humidity and high temperature but cluster 4 represents early noon while cluster 5 represents early night. Cluster 3 has high humidity and low temperature while cluster 5 has low humidity and high temperature but both represent early night.

```{r gmm-hist-comparisons}
clusters_real_data %>%
  filter(Cluster == 4 | Cluster == 5) %>%
  ggplot(., aes(x=Hour, fill=Time)) +
  geom_bar() +
  labs(title="Clusters 4 & 5 Histogram by Hour", y="# of Observations")

clusters_real_data %>%
  filter(Cluster == 4 | Cluster == 5) %>%
  ggplot(., aes(x=Height, fill=Time)) +
  geom_histogram(position="fill", binwidth=5) +
  labs(title="Clusters 4 & 5 Histogram by Height", y="Proportion of Observations")

clusters_real_data %>%
  filter(Cluster == 3 | Cluster == 5) %>%
  ggplot(., aes(x=Day, fill=Cluster)) +
  geom_bar(position="fill") +
  labs(title="Clusters 3 & 5 Histogram by Day", y="Proportion of Observations")

clusters_real_data %>%
  filter(Cluster == 3 | Cluster == 5) %>%
  ggplot(., aes(x=Height, fill=Cluster)) +
  geom_histogram(position="fill", binwidth=5) +
  labs(title="Clusters 3 & 5 Histogram by Height", y="Proportion of Observations")
```

## 5

### a)

get rid of 0s and say why

```{r 5a}
i_par_plot <- data_all %>%
  mutate(hamatop = case_when(
    hamatop == 0 ~ 1,
    T ~ hamatop
  )) %>%
  ggplot(., aes(x=log(hamatop))) +
  geom_histogram(aes(y = (..count..)/sum(..count..)),binwidth=0.5) +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Incident PAR", x="log \u00b5mol/m\u00b2/s", y="Percentage of readings")

r_par_plot <- data_all %>%
  mutate(hamabot = case_when(
    hamabot == 0 ~ 1,
    T ~ hamabot
  )) %>%
  ggplot(., aes(x=log(hamabot))) +
  geom_histogram(aes(y = (..count..)/sum(..count..)),binwidth=0.5) +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Reflected PAR", x="log \u00b5mol/m\u00b2/s", y="Percentage of readings")

grid.arrange(i_par_plot, r_par_plot, nrow=1)
```

### b)

```{r 5b-1}
p_3c_1 <- data_all %>%
  group_by(Height) %>%
  mutate(x1 = median(humid_temp),
         x2 = quantile(humid_temp, probs=0.25),
         x3 = quantile(humid_temp, probs=0.75)) %>%
  distinct(Height, x1, x2, x3) %>%
  ggplot(., aes(x=Height)) +
  geom_point(aes(y=x1, color="Median"), alpha=0.5) +
  geom_line(aes(y=x1), color="red", alpha=0.5) +
  geom_point(aes(y=x2, color="25%"), alpha=0.5) +
  geom_line(aes(y=x2), color="blue", alpha=0.5) +
  geom_point(aes(y=x3, color="75%"), alpha=0.5) +
  geom_line(aes(y=x3), color="green", alpha=0.5) +
  labs(title="Temperature vs Height", x="Height (m)", y="\u00b0C", color="Quantile") +
  scale_color_manual(values=c("Median"="red", "25%"="blue", "75%"="green"))

p_3c_2 <- data_all %>%
  group_by(Height) %>%
  mutate(x1 = median(humid_adj),
         x2 = quantile(humid_adj, probs=0.25),
         x3 = quantile(humid_adj, probs=0.75)) %>%
  distinct(Height, x1, x2, x3) %>%
  ggplot(., aes(x=Height)) +
  geom_point(aes(y=x1, color="Median"), alpha=0.5) +
  geom_line(aes(y=x1), color="red", alpha=0.5) +
  geom_point(aes(y=x2, color="25%"), alpha=0.5) +
  geom_line(aes(y=x2), color="blue", alpha=0.5) +
  geom_point(aes(y=x3, color="75%"), alpha=0.5) +
  geom_line(aes(y=x3), color="green", alpha=0.5) +
  labs(title="Relative Humidity vs Height", x="Height (m)", y="%RH", color="Quantile") +
  scale_color_manual(values=c("Median"="red", "25%"="blue", "75%"="green"))

p_3c_3 <- data_all %>%
  group_by(Height) %>%
  mutate(x1 = median(hamatop),
         x2 = quantile(hamatop, probs=0.25),
         x3 = quantile(hamatop, probs=0.75)) %>%
  distinct(Height, x1, x2, x3) %>%
  ggplot(., aes(x=Height)) +
  geom_point(aes(y=x1, color="Median"), alpha=0.5) +
  geom_line(aes(y=x1), color="red", alpha=0.5) +
  geom_point(aes(y=x2, color="25%"), alpha=0.5) +
  geom_line(aes(y=x2), color="blue", alpha=0.5) +
  geom_point(aes(y=x3, color="75%"), alpha=0.5) +
  geom_line(aes(y=x3), color="green", alpha=0.5) +
  labs(title="Incident PAR vs Height", x="Height (m)", y="\u00b5mol/m\u00b2/s", color="Quantile") +
  scale_color_manual(values=c("Median"="red", "25%"="blue", "75%"="green"))

p_3c_4 <- data_all %>%
  group_by(Height) %>%
  mutate(x1 = median(hamabot),
         x2 = quantile(hamabot, probs=0.25),
         x3 = quantile(hamabot, probs=0.75)) %>%
  distinct(Height, x1, x2, x3) %>%
  ggplot(., aes(x=Height)) +
  geom_point(aes(y=x1, color="Median"), alpha=0.5) +
  geom_line(aes(y=x1), color="red", alpha=0.5) +
  geom_point(aes(y=x2, color="25%"), alpha=0.5) +
  geom_line(aes(y=x2), color="blue", alpha=0.5) +
  geom_point(aes(y=x3, color="75%"), alpha=0.5) +
  geom_line(aes(y=x3), color="green", alpha=0.5) +
  labs(title="Reflected PAR vs Height", x="Height (m)", y="\u00b5mol/m\u00b2/s", color="Quantile") +
  scale_color_manual(values=c("Median"="red", "25%"="blue", "75%"="green"))

grid.arrange(p_3c_1, p_3c_2, p_3c_3, p_3c_4, nrow=2, ncol=2)
```

```{r 5b-2}
p_3d_1 <- data_all %>%
  mutate(humid_temp = humid_temp - mean(humid_temp)) %>%
  group_by(Height) %>%
  mutate(x1 = median(humid_temp),
         x2 = quantile(humid_temp, probs=0.25),
         x3 = quantile(humid_temp, probs=0.75)) %>%
  distinct(Height, x1, x2, x3) %>%
  ggplot(., aes(x=Height)) +
  geom_point(aes(y=x1, color="Median"), alpha=0.5) +
  geom_line(aes(y=x1), color="red", alpha=0.5) +
  geom_point(aes(y=x2, color="25%"), alpha=0.5) +
  geom_line(aes(y=x2), color="blue", alpha=0.5) +
  geom_point(aes(y=x3, color="75%"), alpha=0.5) +
  geom_line(aes(y=x3), color="green", alpha=0.5) +
  labs(title="Temperature vs Height", x="Height (m)", y="Mean Centered \u00b0C", color="Quantile") +
  scale_color_manual(values=c("Median"="red", "25%"="blue", "75%"="green"))

p_3d_2 <- data_all %>%
  mutate(humid_adj = humid_adj - mean(humid_adj)) %>%
  group_by(Height) %>%
  mutate(x1 = median(humid_adj),
         x2 = quantile(humid_adj, probs=0.25),
         x3 = quantile(humid_adj, probs=0.75)) %>%
  distinct(Height, x1, x2, x3) %>%
  ggplot(., aes(x=Height)) +
  geom_point(aes(y=x1, color="Median"), alpha=0.5) +
  geom_line(aes(y=x1), color="red", alpha=0.5) +
  geom_point(aes(y=x2, color="25%"), alpha=0.5) +
  geom_line(aes(y=x2), color="blue", alpha=0.5) +
  geom_point(aes(y=x3, color="75%"), alpha=0.5) +
  geom_line(aes(y=x3), color="green", alpha=0.5) +
  labs(title="Relative Humidity vs Height", x="Height (m)", y="Mean Centered %RH", color="Quantile") +
  scale_color_manual(values=c("Median"="red", "25%"="blue", "75%"="green"))

p_3d_3 <- data_all %>%
  mutate(hamatop = hamatop - mean(hamatop)) %>%
  group_by(Height) %>%
  mutate(x1 = median(hamatop),
         x2 = quantile(hamatop, probs=0.25),
         x3 = quantile(hamatop, probs=0.75)) %>%
  distinct(Height, x1, x2, x3) %>%
  ggplot(., aes(x=Height)) +
  geom_point(aes(y=x1, color="Median"), alpha=0.5) +
  geom_line(aes(y=x1), color="red", alpha=0.5) +
  geom_point(aes(y=x2, color="25%"), alpha=0.5) +
  geom_line(aes(y=x2), color="blue", alpha=0.5) +
  geom_point(aes(y=x3, color="75%"), alpha=0.5) +
  geom_line(aes(y=x3), color="green", alpha=0.5) +
  labs(title="Incident PAR vs Height", x="Height (m)", y="Mean Centered \u00b5mol/m\u00b2/s", color="Quantile") +
  scale_color_manual(values=c("Median"="red", "25%"="blue", "75%"="green"))

p_3d_4 <- data_all %>%
  mutate(hamabot = hamabot - mean(hamabot)) %>%
  group_by(Height) %>%
  mutate(x1 = median(hamabot),
         x2 = quantile(hamabot, probs=0.25),
         x3 = quantile(hamabot, probs=0.75)) %>%
  distinct(Height, x1, x2, x3) %>%
  ggplot(., aes(x=Height)) +
  geom_point(aes(y=x1, color="Median"), alpha=0.5) +
  geom_line(aes(y=x1), color="red", alpha=0.5) +
  geom_point(aes(y=x2, color="25%"), alpha=0.5) +
  geom_line(aes(y=x2), color="blue", alpha=0.5) +
  geom_point(aes(y=x3, color="75%"), alpha=0.5) +
  geom_line(aes(y=x3), color="green", alpha=0.5) +
  labs(title="Reflected PAR vs Height", x="Height (m)", y="Mean Centered \u00b5mol/m\u00b2/s", color="Quantile") +
  scale_color_manual(values=c("Median"="red", "25%"="blue", "75%"="green"))

grid.arrange(p_3d_1, p_3d_2, p_3d_3, p_3d_4, nrow=2, ncol=2)
```

### c)

add height to legend labels

```{r 5c-1}
library(lubridate)
data_all %>%
  filter(as.Date(epochDates) == as.Date("2004-05-01")) %>%
  group_by(nodeid) %>%
  mutate(`Mean Temperature` = mean(humid_temp)) %>%
  ungroup() %>%
  distinct(nodeid, `Mean Temperature`) %>%
  arrange(., desc(`Mean Temperature`)) %>%
  slice(1:3)

data_all %>%
  filter(as.Date(epochDates) == as.Date("2004-05-01")) %>%
  group_by(nodeid) %>%
  mutate(`Mean Temperature` = mean(humid_temp)) %>%
  ungroup() %>%
  distinct(nodeid, `Mean Temperature`) %>%
  arrange(., `Mean Temperature`) %>%
  slice(1:3)

data_all %>%
  filter(as.Date(epochDates) == as.Date("2004-05-01")) %>%
  group_by(nodeid) %>%
  mutate(`Mean Temperature` = mean(humid_temp)) %>%
  ungroup() %>%
  distinct(nodeid, `Mean Temperature`) %>%
  arrange(., `Mean Temperature`) %>%
  slice(26:28)

data_all %>%
  filter((as.Date(epochDates) == as.Date("2004-05-01")) & (nodeid == 118 | nodeid == 20 | nodeid == 11)) %>%
  mutate(nodeid = factor(nodeid),
         Time = format(epochDates, "%H:%M")) %>%
  ggplot(., aes(x=Time)) +
  geom_point(aes(y=humid_temp, color=nodeid)) +
  geom_line(aes(y=humid_temp, group=nodeid, color=nodeid)) +
  scale_x_discrete(breaks=c("00:00", "06:00", "12:00", "18:00", "23:30")) +
  labs(title="May 1 Temperature vs Time", y="Temperature (\u00b0C)", color="Node id")
```

```{r 5c-2}
library(lubridate)
data_all %>%
  filter(as.Date(epochDates) == as.Date("2004-05-01")) %>%
  group_by(nodeid) %>%
  mutate(`Mean Humidity` = mean(humid_adj)) %>%
  ungroup() %>%
  distinct(nodeid, `Mean Humidity`) %>%
  arrange(., desc(`Mean Humidity`)) %>%
  slice(1:3)

data_all %>%
  filter(as.Date(epochDates) == as.Date("2004-05-01")) %>%
  group_by(nodeid) %>%
  mutate(`Mean Humidity` = mean(humid_adj)) %>%
  ungroup() %>%
  distinct(nodeid, `Mean Humidity`) %>%
  arrange(., `Mean Humidity`) %>%
  slice(1:3)

data_all %>%
  filter(as.Date(epochDates) == as.Date("2004-05-01")) %>%
  group_by(nodeid) %>%
  mutate(`Mean Humidity` = mean(humid_adj)) %>%
  ungroup() %>%
  distinct(nodeid, `Mean Humidity`) %>%
  arrange(., `Mean Humidity`) %>%
  slice(26:28)

data_all %>%
  filter((as.Date(epochDates) == as.Date("2004-05-01")) & (nodeid == 20 | nodeid == 17 | nodeid == 115)) %>%
  mutate(nodeid = factor(nodeid),
         Time = format(epochDates, "%H:%M")) %>%
  ggplot(., aes(x=Time)) +
  geom_point(aes(y=humid_adj, color=nodeid)) +
  geom_line(aes(y=humid_adj, group=nodeid, color=nodeid)) +
  scale_x_discrete(breaks=c("00:00", "06:00", "12:00", "18:00", "23:30")) +
  labs(title="May 1 Relative Humidity vs Time", y="%RH", color="Node id")
```

### d)

114980 network observations

putting net and log on same graphs for 1,2,3

```{r 5d-net-merges}
data_net <- merge(data_net, dates, by = "epoch", all.x = T, suffixes = c(".net", ".dates"))
data_net <- merge(data_net, locations, by = "nodeid", all.x = T, suffixes = c(".net", ".locations"))
```

```{r 5d-net-yields-dates}
net_dates_yield <- c()
dates_range <- c()
nodes_per <- c()
for (date in as.list(unique(as.Date(dates$epochDates)))) {
  date_yield <- data_net %>%
    filter(as.Date(epochDates) == as.Date(date) & !is.na(voltage)) %>%
    distinct(nodeid) %>%
    nrow()
  net_dates_yield <- c(net_dates_yield, date_yield/nrow(locations))
  
  date_nodes <- data_net %>%
    filter(as.Date(epochDates) == as.Date(date) & !is.na(voltage)) %>%
    distinct(nodeid) %>%
    pull()
  d <- rep(date, length(date_nodes))
  dates_range <- c(dates_range, d)
  nodes_per <- c(nodes_per, date_nodes)
}

yields_dates <- dates %>%
  distinct(as.Date(epochDates)) %>%
  cbind(., net_dates_yield)
names(yields_dates) <- c("Day", "Yield")

yields_dates %>%
  ggplot(., aes(x=100*Yield)) +
  geom_bar(aes(y = 100*(..count..)/sum(..count..))) +
  labs(title="Per Day Yields", x="Yield %", y="% of Days")

yields_dates %>%
  ggplot(., aes(x=Day, y=100*Yield)) +
  geom_point() +
  geom_line() +
  labs(title="Per Day Yields", y="Yield %")

data.frame(cbind(dates_range, nodes_per)) %>%
  merge(., locations, by.x="nodes_per", by.y="nodeid", all.x=T) %>%
  mutate(Height = as.numeric(Height),
         dates_range = as.Date(dates_range, origin="1970-01-01")) %>%
  ggplot(., aes(x=dates_range, y=Height)) +
  geom_point()
```

```{r 5d-net-yields-2}
net_nodes_yield <- c()
for (node_id in locations$nodeid) {
  node_yield <- data_net %>%
    filter(nodeid == node_id & !is.na(voltage)) %>%
    distinct(as.Date(epochDates)) %>%
    nrow()
  net_nodes_yield <- c(net_nodes_yield, node_yield/length(unique(as.Date(dates$epochDates))))
}

yields_nodes <- locations %>%
  mutate(Height = as.numeric(Height)) %>%
  distinct(nodeid, Height) %>%
  cbind(., net_nodes_yield)
names(yields_nodes) <- c("Node ID", "Height", "Yield")

yields_nodes %>%
  ggplot(., aes(x=100*Yield)) +
  geom_bar(aes(y = 100*(..count..)/sum(..count..))) +
  labs(title="Per Node Yields", x="Yield %", y="% of Nodes")

yields_nodes %>%
  ggplot(., aes(x=100*Yield, y=Height)) +
  geom_point() +
  labs(title="Per Node Yields", x="Yield %", y="Node Height (m)")
```
